---
title: "Video evaluation via SoSci"
author: "Irene Sophia Plank"
date: "2022-10-05"
output: html_document
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}

knitr::opts_chunk$set(echo = TRUE)
if (file.exists('/home/iplank/Insync/10planki@gmail.com/Google Drive/NEVIA/study_PRED/stim_evaluation/')){
  knitr::opts_knit$set(root.dir = '/home/iplank/Insync/10planki@gmail.com/Google Drive/NEVIA/study_PRED/stim_evaluation/')
} else if (file.exists('C:/Users/Oswin/Google Drive/NEVIA/study_PRED/stim_evaluation')){
  knitr::opts_knit$set(root.dir = 'C:/Users/Oswin/Google Drive/NEVIA/study_PRED/stim_evaluation')
}

library(knitr)           # kable
library(tidyverse)       # tibble stuff
library(data.table)      # melt / dcast
library(lmerTest)        # lmer function
library(jtools)          # summ(model, digits=3, confint=T)
library(sjPlot)          # tab_model
library(JWileymisc)      # modelPerformance and modelTest
library(multilevelTools) # to use modelTest and modelPerformance with lmer
library(car)             # contr.Sum function
library(ggplot2)         # plots
library(bayestestR)      # Bayesian model comparison
library(performance)     # check_model and compare_performance
library(effectsize)      # allows to interpret effect sizes
library(ggpubr)          # ggarrange

```

<style type="text/css">
.main-container {
  max-width: 1100px;
  margin-left: auto;
  margin-right: auto;
}
</style>

# Setup

The data is loaded and relevant columns are selected. Then, it is transformed to the long format and the Video names are added. In the end, all information is contained in one data frame that is in the long format. 

```{r load_data, include = FALSE}

# Load data ---------------------------------------------------------------

df_fem      = read_delim('data_emopred_fer_2022-10-04_13-31.csv', delim = ";", show_col_types = FALSE)
df_mal      = read_delim('data_emopred_fer_2_2022-10-04_13-32.csv', delim = ";", show_col_types = FALSE)

# make column names for comments always starting with CO and not C0
colnames(df_fem) = sub("C0", "CO", colnames(df_fem))
colnames(df_mal) = sub("C0", "CO", colnames(df_mal))

# select only finished questionnaires and relevant columns
df_fem_sel  = df_fem %>% filter(FINISHED == 1) %>% select(matches("CASE|CO|NA|RG.._"))
df_mal_sel  = df_mal %>% filter(FINISHED == 1) %>% select(matches("CASE|CO|NA|RG.._"))

# convert to long format
df_fem_long = melt(data = df_fem_sel,
                   id.vars= "CASE",
                   variable.name = "VARIABLE",
                   value.name = "VALUE")
df_mal_long = melt(data = df_mal_sel,
                   id.vars= "CASE",
                   variable.name = "VARIABLE",
                   value.name = "VALUE")

# divide the info into columns
df_fem_long$QUE = substr(df_fem_long$VARIABLE, 1, 2)
df_mal_long$QUE = substr(df_mal_long$VARIABLE, 1, 2)
df_fem_long$VID = substr(df_fem_long$VARIABLE, 3, 4)
df_mal_long$VID = substr(df_mal_long$VARIABLE, 3, 4)
df_fem_long$NUM = substr(df_fem_long$VARIABLE, 6, 7)
df_mal_long$NUM = substr(df_mal_long$VARIABLE, 6, 7)

# recode the VID column
df_fem_long = mutate(df_fem_long, VID = fct_recode(VID, 
                      "AF01AFS_gimped" = "01", "AF01ANS_gimped" = "02", "AF01HAS_gimped" = "03", "AF01SAS_gimped" = "04",
                      "AF02AFS_gimped" = "05", "AF02SAS_gimped" = "06", "AF05AFS_gimped" = "07", "AF05ANS_gimped" = "08",
                      "AF05HAS_gimped" = "09", "AF05SAS_gimped" = "10", "AF06AFS_gimped" = "11", "AF06ANS_gimped" = "12",
                      "AF06HAS_gimped" = "13", "AF06SAS_gimped" = "14", "AF09HAS_gimped" = "15", "AF09SAS_gimped" = "16",
                      "AF11AFS_gimped" = "17", "AF11ANS_gimped" = "18", "AF11SAS_gimped" = "19", "AF14AFS_gimped" = "20",
                      "AF14ANS_gimped" = "21", "AF14HAS_gimped" = "22", "AF14SAS_gimped" = "23", "AF29AFS_gimped" = "24",
                      "AF29ANS_gimped" = "25", "AF29HAS_gimped" = "26", "AF29SAS_gimped" = "27", "AF32AFS_gimped" = "28",
                      "AF32ANS_gimped" = "29", "AF32HAS_gimped" = "30", "AF32SAS_gimped" = "31", "BF02ANS_gimped" = "32",
                      "BF02HAS_gimped" = "33", "BF07AFS_gimped" = "34", "BF07ANS_gimped" = "35", "BF07HAS_gimped" = "36",
                      "BF07SAS_gimped" = "37", "BF08AFS_gimped" = "38", "BF08ANS_gimped" = "39", "BF08SAS_gimped" = "40",
                      "BF08HAS_gimped" = "41", "AF09ANS_gimped" = "42", "AF11HAS_gimped" = "43", "AF16AFS_gimped" = "44",
                      "AF16HAS_gimped" = "45", "AF16SAS_gimped" = "46", "AF19AFS_gimped" = "47", "AF19ANS_gimped" = "48",
                      "AF19HAS_gimped" = "49", "AF19SAS_gimped" = "50", "AF20AFS_gimped" = "51", "AF20ANS_gimped" = "52",
                      "AF20HAS_gimped" = "53", "AF20SAS_gimped" = "54", "AF24AFS_gimped" = "55", "AF24ANS_gimped" = "56",
                      "AF24HAS_gimped" = "57", "AF24SAS_gimped" = "58", "AF27AFS_gimped" = "59", "AF27ANS_gimped" = "60",
                      "AF27HAS_gimped" = "61", "AF27SAS_gimped" = "62", "BF16ANS_gimped" = "63", "AF09AFS_gimped" = "64",
                      ))  

df_mal_long = mutate(df_mal_long, VID = fct_recode(VID, 
                      "AM01ANS_gimped" = "01", "AM01HAS_gimped" = "02", "AM01SAS_gimped" = "03", "AM02AFS_gimped" = "04",
                      "AM02ANS_gimped" = "05", "AM02HAS_gimped" = "06", "AM02SAS_gimped" = "07", "AM03AFS_gimped" = "08",
                      "AM03ANS_gimped" = "09", "AM03HAS_gimped" = "10", "AM03SAS_gimped" = "11", "AM05AFS_gimped" = "12",
                      "AM05ANS_gimped" = "13", "AM05HAS_gimped" = "14", "AM05SAS_gimped" = "15", "AM06AFS_gimped" = "16",
                      "AM06ANS_gimped" = "17", "AM06HAS_gimped" = "18", "AM06SAS_gimped" = "19", "AM07AFS_gimped" = "20",
                      "AM07ANS_gimped" = "21", "AM07HAS_gimped" = "22", "AM07SAS_gimped" = "23", "AM08AFS_gimped" = "24",
                      "AM08ANS_gimped" = "25", "AM08HAS_gimped" = "26", "AM08SAS_gimped" = "27", "AM09AFS_gimped" = "28",
                      "AM09ANS_gimped" = "29", "AM09HAS_gimped" = "30", "AM09SAS_gimped" = "31", "AM10AFS_gimped" = "32",
                      "AM10ANS_gimped" = "33", "AM10HAS_gimped" = "34", "AM10SAS_gimped" = "35", "AM01AFS_gimped" = "36",
                      "AM11AFS_gimped" = "37", "AM11ANS_gimped" = "38", "AM11HAS_gimped" = "39", "AM11SAS_gimped" = "40",
                      "AM13AFS_gimped" = "41", "AM13ANS_gimped" = "42", "AM13HAS_gimped" = "43", "AM13SAS_gimped" = "44",
                      "AM14AFS_gimped" = "45", "AM14ANS_gimped" = "46", "AM14HAS_gimped" = "47", "AM14SAS_gimped" = "48",
                      "BM23AFS_gimped" = "49", "BM23ANS_gimped" = "50", "BM23HAS_gimped" = "51", "BM23SAS_gimped" = "52",
                      "BM27AFS_gimped" = "53", "BM27ANS_gimped" = "54", "BM27HAS_gimped" = "55", "BM27SAS_gimped" = "56",
                      "BM28AFS_gimped" = "57", "BM28ANS_gimped" = "58", "BM28HAS_gimped" = "59", "BM28SAS_gimped" = "60",
                      "BM31AFS_gimped" = "61", "BM31ANS_gimped" = "62", "BM31HAS_gimped" = "63", "BM31SAS_gimped" = "64",
                      ))  

# split into question types
df_mal_CO = df_mal_long %>% filter(QUE == "CO") %>% select(-c(NUM, QUE, VARIABLE)) %>% rename("COM" = "VALUE")
df_mal_NA = df_mal_long %>% filter(QUE == "NA") %>% select(-c(NUM, QUE)) %>% rename("NAT" = "VALUE") %>% mutate(NAT = as.numeric(NAT))
df_mal_RG = df_mal_long %>% filter(QUE == "RG") %>% select(-c(QUE, VARIABLE)) %>% rename("REG" = "VALUE") %>%
  mutate(NUM = fct_recode(NUM, 
                          "REG_brow" = "01",
                          "REG_eyes" = "02",
                          "REG_nose" = "03",
                          "REG_mouth" = "04",
                          "REG_other" = "05",
                          ),
         REG = as.numeric(REG) - 1, 
         EMO = case_when(
           str_detect(VID, "AFS") ~ "AF",
           str_detect(VID, "HAS") ~ "HA",
           str_detect(VID, "SAS") ~ "SA",
           str_detect(VID, "ANS") ~ "AN"
           ))

df_fem_CO = df_fem_long %>% filter(QUE == "CO") %>% select(-c(NUM, QUE, VARIABLE)) %>% rename("COM" = "VALUE")
df_fem_NA = df_fem_long %>% filter(QUE == "NA") %>% select(-c(NUM, QUE)) %>% rename("NAT" = "VALUE") %>% mutate(NAT = as.numeric(NAT))
df_fem_RG = df_fem_long %>% filter(QUE == "RG") %>% select(-c(QUE, VARIABLE)) %>% rename("REG" = "VALUE") %>%
  mutate(NUM = fct_recode(NUM, 
                          "REG_brow" = "01",
                          "REG_eyes" = "02",
                          "REG_nose" = "03",
                          "REG_mouth" = "04",
                          "REG_other" = "05",
                          ),
         REG = as.numeric(REG) - 1, 
         EMO = case_when(
           str_detect(VID, "AFS") ~ "AF",
           str_detect(VID, "HAS") ~ "HA",
           str_detect(VID, "SAS") ~ "SA",
           str_detect(VID, "ANS") ~ "AN"
           ))

# convert region into mixed format
df_mal_RG = dcast(df_mal_RG, CASE + VID + EMO ~ NUM, value.var = "REG")
df_fem_RG = dcast(df_fem_RG, CASE + VID + EMO ~ NUM, value.var = "REG")

# merge question types again
df_mal = merge(df_mal_NA,df_mal_RG)
df_mal = merge(df_mal,df_mal_CO, all.x = T)
df_fem = merge(df_fem_NA,df_fem_RG)
df_fem = merge(df_fem,df_fem_CO, all.x = T)

# combine male and female videos
df = rbind(df_fem, df_mal)

# add info about the survey to the case ID
df$CASE = paste(substr(df$VID,2,2), df$CASE, sep = "")

# load list of ratings to exclude based on comments (glitch)
if (file.exists("df_COM.csv")) {
  df_COMexc = read_csv("df_COM.csv", show_col_types = FALSE)
} 
df = merge(df,df_COMexc[,c("CASE","VID","COM_exc")], all.x = T)
df[is.na(df$COM_exc),"COM_exc"] = 0
df[df$COM_exc == 1,"NAT"] = NA

# add a gender and face ID column
df$GEN = substr(df$VID, 2, 2)
df$FID = substr(df$VID, 2, 4)

# check out all NAs and see if they are in the original data
test_M = df %>% filter(is.na(NAT) & is.na(COM) & GEN == "M") %>% select(c(CASE, VARIABLE)) # yes, these are also missing in original data 
test_F = df %>% filter(is.na(NAT) & is.na(COM) & GEN == "F") %>% select(c(CASE, VARIABLE)) # yes, these are also missing in original data

# drop all unnecessary variables
df = df %>% select(-c(COM_exc, VARIABLE))
rm(list = setdiff(ls(), c("df")))


```
# Choosing videos for the FER task

For the FER task, we want to choose the most natural video groups. Two things are important: that the videos are consistently rated as natural (median above 50) and that this is not due to a few very high ratings. First, outliers are removed that are lower than or equal to the 25% quantile minus 1.5 times the interquartile range. Then, face IDs are chosen who have the highest minima and a median per emotion of at least 50. We chose the median instead of the mean because the distributions are not at all normally distributed. The comments concerning the selected videos are saved to further improve them. 

``` {r sel_FER}

# get rid of extreme outliers for each video
df_fil = df %>% group_by(VID, FID, GEN) %>%
  filter(NAT > quantile(NAT, na.rm = T)[2] - 1.5*IQR(NAT, na.rm = T))

# aggregate per face ID and emotion
df_agg = df_fil %>% group_by(VID, GEN, EMO) %>%
  summarise(
    NAT_med = median(NAT, na.rm = T),
    NAT_min = min(NAT, na.rm = T), 
    NAT_val = n()
  ) %>% filter(NAT_med >= 50 & NAT_val > 10)

# choose the videos with the highest minima
n = 8 # number of videos per gender and emotion
vid_sel = df_agg %>%
  arrange(desc(NAT_min)) %>% 
  group_by(GEN, EMO) %>%
  slice(1:n) %>%
  select(GEN, EMO, VID)

# filter the full data frame to only include selected videos
df_FER_sel = df_fil %>% filter(VID %in% vid_sel$VID)

# aggregate info for the selected videos
df_FER_sel_agg = df_FER_sel %>% group_by(FID, VID, EMO, GEN) %>%
  summarise(
    NAT_avg = mean(NAT, na.rm = T),
    NAT_ste = sd(NAT, na.rm = T) / sqrt(n()), 
    NAT_med = median(NAT, na.rm = T),
    NAT_min = min(NAT, na.rm = T)
  ) %>%
  arrange(FID, EMO, VID)
kable(df_FER_sel_agg)

# # get all comments 
# df_FER_sel_COM = df_FER_sel %>% filter(!is.na(COM)) %>% select(GEN, FID, VID, COM) %>%
#   arrange(FID, VID)
# kable(df_FER_sel_COM)

# drop all unnecessary variables
rm(list = setdiff(ls(), c("df", "df_FER_sel", "df_FER_sel_agg", "df_FER_sel_COM")))


```

## Visualisation

``` {r sel_FER_vis}

# set colour scheme
cls = c("#007e36", "#86eb9a") 
names(cls) = c("M", "F")

# create boxplots
ggplot(df_FER_sel, aes(x = EMO, y = NAT, fill = GEN)) +
  geom_boxplot() +
  labs (x = "emotion", y = "natural") + 
  ggtitle("Average natural rating") + 
  theme_classic() + scale_fill_manual(values = cls) + 
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))


```

## Gender and emotion effects

A linear mixed model is computed to assess whether the four emotions of the selected videos are comparable in terms of natural ratings. Here, we use sum coding for the emotion and compare each emotion to the grand mean of all of them. This is explained more clearly on [CrossValidated](https://stats.stackexchange.com/questions/320740/how-to-interpret-sum-contrast-in-regression-lmm): 

> <font size="2">`contr.sum` makes sure all the contrasts sum to zero so that the "intercept" term is the grand mean. The effects are summarized with coefficients representing the number of factor levels ($k$) minus 1. In this case, the last level is dropped. The coefficients represent a mean difference from the "grand mean" for the first $k−1$ factor levels.

> If you fit `lm(y ~ x - 1)` where $x$ has the sum contrast, then the model coefficients are the stratum specific means for each level of $x$, but the inference tests a mean difference from 0 for each level of $x$. This is useless unless $x$ [!sic?] is centered. I suggest you fit this model if you wish to simultaneously test the 3 parameters' difference from the grand mean. Technically speaking, this is overkill since such a model is overspecified, but that isn't inherently a bad thing.
</font size>


``` {r lmm_FER}

# center naturalnes
df_FER_sel$NAT = scale(df_FER_sel$NAT, scale = F)

# set factors correctly
df_FER_sel$EMO = as.factor(df_FER_sel$EMO)
contrasts(df_FER_sel$EMO)  = contr.Sum(levels(df_FER_sel$EMO))

# model -> each level of EMO separately? why?
m1 = lmer(formula = NAT ~ EMO - 1 + # fixed effects
                         (1 | CASE) + (1 | VID), # random effects
                         data = df_FER_sel, REML = T)

tab_model(m1, show.df = F, show.stat = T, show.ci = F, linebreak = F, wrap.labels = 80)

set_theme(base = theme_classic())

plot_model(m1)

```

# Choosing videos for the PAL task

Since in the PAL task, we do not use the videos themselves but three stills of different emotion strengths, we can simply choose the videos with the best median natural ratings. However, to keep it consistent with the Lawson et al. (2017) paper, we will use six identities which are shown in each emotion. 

``` {r sel_PAL}

# get rid of extreme outliers for each video
df_fil = df %>% group_by(VID, FID, GEN) %>%
  filter(NAT > quantile(NAT, na.rm = T)[2] - 1.5*IQR(NAT, na.rm = T))

# aggregate per face ID and emotion
df_agg = df_fil %>% group_by(VID, FID, GEN, EMO) %>%
  summarise(
    NAT_med = median(NAT, na.rm = T)
  ) %>% filter(VID %in% df_FER_sel_agg$VID & (EMO == "AF" | EMO == "HA"))

# find out of which IDs both emotions are left
fids = df_agg$FID[duplicated(df_agg$FID)]
df_agg = df_agg %>% filter(FID %in% fids)

# choose the videos with the highest median
n = 3 # number of videos per gender and emotion
fid_sel = df_agg %>% group_by(GEN, FID) %>% 
  summarise(NAT_med = mean(NAT_med)) %>%
  arrange(desc(NAT_med)) %>% 
  group_by(GEN) %>%
  slice(1:n) %>%
  select(GEN, FID)

# filter the full data frame to only include selected videos
df_PAL_sel = df_fil %>% filter(FID %in% fid_sel$FID & (EMO == "AF" | EMO == "HA"))

# aggregate info for the selected videos
df_PAL_sel_agg = df_PAL_sel %>% group_by(FID, VID, EMO, GEN) %>%
  summarise(
    NAT_avg = mean(NAT, na.rm = T),
    NAT_ste = sd(NAT, na.rm = T) / sqrt(n()), 
    NAT_med = median(NAT, na.rm = T),
    NAT_min = min(NAT, na.rm = T)
  ) %>%
  arrange(FID, EMO, VID)
kable(df_PAL_sel_agg)

# # get all comments 
# df_PAL_sel_COM = df_PAL_sel %>% filter(!is.na(COM)) %>% select(GEN, FID, VID, COM) %>%
#   arrange(FID, VID)
# kable(df_PAL_sel_COM)

```


## Visualisation

``` {r sel_PAL_vis}

# create boxplots
ggplot(df_PAL_sel, aes(x = EMO, y = NAT, fill = GEN)) +
  geom_boxplot() +
  labs (x = "emotion", y = "natural") + 
  ggtitle("Average natural rating") + 
  theme_classic() + scale_fill_manual(values = cls) + 
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))


```

## Gender and emotion effects

The same linear mixed model is computed to assess whether the two emotions of the selected videos are comparable in terms of natural ratings. 


``` {r lmm_PAL}

# center naturalnes
df_PAL_sel$NAT = scale(df_PAL_sel$NAT, scale = F)

# set factors correctly
df_PAL_sel$EMO = as.factor(df_PAL_sel$EMO)
contrasts(df_PAL_sel$EMO)  = contr.Sum(levels(df_PAL_sel$EMO))

# model -> each level of EMO separately? why?
m2 = lmer(formula = NAT ~ EMO - 1 + # fixed effects
                         (1 | CASE) + (1 | VID), # random effects
                         data = df_PAL_sel, REML = T)

tab_model(m2, show.df = F, show.stat = T, show.ci = F, linebreak = F, wrap.labels = 80)

plot_model(m2)

```