---
title: "S1: behavioural, conventional analysis with brms"
author: "I. S. Plank"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
fontsize: 10pt
#geometry: margin=0.5in
---
  
```{r settings, include=FALSE}

knitr::opts_chunk$set(echo = T, warning = F, message = F, fig.align = 'center', fig.width = 7.4)
ls.packages = c("knitr",# kable
    "ggplot2",          # plots
    "brms",             # Bayesian lmms
    "bridgesampling",   # bridge_sampler
    "tidyverse",        # tibble stuff
    "ggpubr",           # ggarrange
    "ggrain",           # geom_rain
    "bayesplot",        # plots for posterior predictive checks
    "SBC",              # plots for checking computational faithfulness
    "rstatix",          # anova
    "officer",          # read_docx
    "effectsize",       # interpret_bf
    "BayesFactor", 
    "bayestestR"        # equivalence_test
)

lapply(ls.packages, library, character.only=TRUE)

# options for summarise
options(dplyr.summarise.inform = FALSE)

# set cores
options(mc.cores = parallel::detectCores())

# set options for SBC
use_cmdstanr = getOption("SBC.vignettes_cmdstanr", TRUE) # Set to false to use rstan instead
options(brms.backend = "cmdstanr")

# using parallel processing
library(future)
plan(multisession)

# Setup caching of results
brms_dir = "./_brms_models"
if(!dir.exists(brms_dir)) {
  dir.create(brms_dir)
}

# load helper functions
source('../helper/createMs.R')
source('../helper/createDFdemo.R')

# graph settings 
c_light = "#a9afb2"; c_light_highlight = "#8ea5b2"; c_mid = "#6b98b2" 
c_mid_highlight = "#3585b2"; c_dark = "#0072b2"; c_dark_highlight = "#0058b2" 
c_green = "#009E73"
c_dark_green = "#006f50"
sz = 1
a = 0.5

# custom colour palette
col.grp = c("#FFC107", "#004D40", "#1E88E5")
col.cnd = c("#5D3A9B", "#E66100")

# confidence interval functions
lower_ci = function(var) {
  unname(quantile(var, probs = 0.025))
}
upper_ci = function(var) {
  unname(quantile(var, probs = 0.975))
}

# calculate visual angle: px.x should be a vector of c(width, height) in pixel
vis_ang = function(px.x) {
  
  # infos on the setup
  mm.w = 344    # monitor width in mm
  mm.h = 215    # monitor height in mm
  px.w = 2650   # monitor resolution: width
  px.h = 1600   # monitor resolution: height
  dist = 570    # viewing distance
  
  # convert width from screen pixel size to mm and take half
  mm.x    = (px.x[1] / (px.w/mm.w))/2
  
  # convert height from screen pixel size to mm and take half
  mm.x[2] = (px.x[2] / (px.h/mm.h))/2
  
  # calculate angles
  rad.alpha    = atan(mm.x[1]/dist)
  rad.alpha[2] = atan(mm.x[2]/dist)
  
  # convert to degrees and double again
  deg.alpha    = (rad.alpha[1] / (pi/180)) * 2
  deg.alpha[2] = (rad.alpha[2] / (pi/180)) * 2
  
  return(deg.alpha)
  
}

```

<style type="text/css">
  .main-container {
    max-width: 1100px;
    margin-left: auto;
    margin-right: auto;
  }
</style>
  
# Task info and setup
  
This R Markdown script analyses data from the PAL (probabilistic associative learning) task of the EMBA project. It focuses on the comparison of adults with ADHD (with or without comorbid ASD) and adults without psychiatric diagnoses. The data was preprocessed before being read into this script. 

In this task, participants view faces for which they have to judge whether the portrayed emotion is positive or negative. The faces are preceded by a tone which is predictive of the valence of the following face. The visual angle of the faces was `r round(vis_ang(c(265,350))[1],2)` degrees wide and `r round(vis_ang(c(265,350))[2],2)` degrees high.

## Plot the development of associations across task

```{r plot_scheme, fig.height=4}

# load the trial order
df.trl = read_csv('../data/PAL_scheme.csv', show_col_types = F) %>%
  mutate(
    association = ut*100, 
    prob_HP = prob_HP*100
  )

# create rectangles to colour the phases: prevolatile, volatile and postvolatile
rects = data.frame(xstart = c(-Inf, 72, 264), 
                   xend = c(72, 264, Inf), 
                   col = c("a", "b", "c"))
  
ggplot() +
  geom_rect(data = rects, aes(xmin = xstart, xmax = xend, 
                              ymin = -Inf, ymax = Inf, 
                              fill = col), alpha = 0.4) +
  scale_fill_manual(values=c(c_mid_highlight, c_green, c_dark_highlight), 
                    guide = "none") +
  geom_jitter(data = df.trl, 
              aes(x = trl, y = association, shape = expected),
              colour = "black", alpha = 0.8, size = 1, width = 0, height = 4) +
  geom_line(data = df.trl, 
            aes(x = trl, y = prob_HP),
            colour = "black") +
  scale_shape_manual(values = c(8, 1, 3)) +
  theme_bw() + 
  scale_y_continuous(breaks=c(16.7, 83.3)) +
  labs(x = "trial", y = "percent", 
       title = "High tone > positive, low tone > negative emotion") +
  annotate("text", x = 36, y = 125, label= "prevolatile", size = 4) +
  annotate("text", x = 168, y = 125, label= "volatile", size = 4) +
  annotate("text", x = 301, y = 125, label= "postvolatile", size = 4) +
  theme(legend.position = c(0.9, 0.5), legend.title = element_blank(), 
        legend.background = element_rect(fill = alpha("white", 0)),
        legend.key = element_rect(fill = alpha("white", 0)),
        plot.title = element_text(hjust = 0.5), text = element_text(size = 12))

ggsave("plots/PAL_scheme.png", 
       units = "cm",
       width = 27,
       height = 9)


```

## Some general settings

```{r set}

# number of simulations
nsim = 250

# set number of iterations and warmup for models
iter = 6000
warm = 1500

# set the seed
set.seed(2468)

```

## Package versions

The following packages are used in this RMarkdown file: 
  
```{r lib_versions, echo=F}

print(R.Version()$version.string)

for (package in ls.packages) {
  print(sprintf("%s version %s", package, packageVersion(package)))
}

```

## Analysis guidelines 

We planned to determine the group-level effect subjects following Barr (2013). For each model, experiment specific priors were set based on previous literature or the task (see comments in the code).

We performed prior predictive checks as proposed in Schad, Betancourt and Vasishth (2020) using the SBC package before running this analysis (see [!LINK]). We performed these data-naive evaluations based on the comparison of three groups. To do so, we create `r nsim` simulated datasets where parameters are simulated from the priors. These parameters are used to create one fake dataset. Both the true underlying parameters and the simulated discrimination values are saved. Then, we create graphs showing the prior predictive distribution of the simulated discrimination threshold to check whether our priors fit our general expectations about the data. Next, we perform checks of computational faithfulness and model sensitivity as proposed by Schad, Betancourt and Vasishth (2020) and implemented in the SBC package. We create models for each of the simulated datasets. Last, we calculate performance metrics for each of these models, focusing on the population-level parameters. 

## Preparation

First, we load the data and combine it with demographic information including the diagnostic status of the subjects. Then, all predictors are set to sum contrasts. 

```{r prep}


# check if the data file exists, if yes load it:
if (!file.exists("../data/PAL-ADHD_data.RData")) {
  
  path = "/home/emba/Documents/EMBA"

  # get demo info for subjects
  df.sub = read_csv(file.path(path, "CentraXX", "EMBA_centraXX.csv"), 
                    show_col_types = F) %>%
    mutate(
      diagnosis = as.factor(recode(diagnosis, "CTR" = "COMP"))
    )
  
  # set the data path
  dt.path  = file.path(path, "BVET")
  dt.explo = file.path(path, "BVET-explo")
  
  # load the preprocessed data: df.tsk
  df.tsk = rbind(readRDS(file.path(dt.path,  "PAL_tsk.rds")),
                 readRDS(file.path(dt.explo, "PAL_tsk.rds")))
  
  # load excluded participants (accuracy < 2/3)
  exc = c(scan(file.path(dt.path, 'PAL_exc.txt'),  what="character", sep=NULL),
          scan(file.path(dt.explo, 'PAL_exc.txt'), what="character", sep=NULL))
  
  # merge with group and focus on comparison and autistic adults
  df.tsk = merge(df.sub %>% select(subID, diagnosis), 
                 df.tsk, all.y = T) %>%
    filter(diagnosis != "ASD" & !(subID %in% exc)) %>%
    droplevels() %>%
    mutate_if(is.character, as.factor)
  
  # only keep participants included in the study in the subject data frame
  df.sub = merge(df.tsk %>% select(subID) %>% distinct(), df.sub, all.x = T) %>%
    droplevels()

  # anonymise the data based on predetermined file > same for all data
  df.recode = read_csv(file.path(path, "BVET", "PID_anonymisation.csv"))
  recode = as.character(df.recode$subID)
  names(recode) = df.recode$PID
  df.tsk$subID = str_replace_all(df.tsk$subID, recode)
  
  # add information on ADHD medication
  df.tsk = merge(df.tsk, df.recode %>% select(subID, adhd.meds, adhd.meds.bin))
  
  # save the data for use with the EMBA_HGF toolbox
  df.tsk %>% select(subID, trl, diagnosis, ut, difficulty, acc, rt, EDT, adhd.meds.bin) %>%
    ungroup() %>%
    arrange(subID, trl) %>%
    write_csv(., file = "../data/PAL-ADHD_data.csv")
  
  # print gender frequencies and compare them across groups
  tb.gen = xtabs(~ gender + diagnosis, data = df.sub)
  ct.full = contingencyTableBF(tb.gen, 
                               sampleType = "indepMulti", 
                               fixedMargin = "cols")
  
  # get all self-descriptions for DAN group
  gen.desc = unique(tolower(df.sub[df.sub$gender == "dan",]$gender_desc))
  
  # create a demographics overview
  ls.demo = createDFdemo(df.sub, 
                         c("age", "edu", "BDI_total", "ASRS_total", "RAADS_total", "TAS_total", "iq"),
                         "diagnosis")
  
  # save it all
  save(df.tsk, ls.demo, ct.full, gen.desc, tb.gen, 
       file = "../data/PAL-ADHD_data.RData")
  
} else {
  
  load("../data/PAL-ADHD_data.RData")
  
}

# print the outcome of the two contingency tables for comparison: all participants
ct.full@bayesFactor
# print gender and diagnosis distribution
kable(tb.gen)

# print medication distribution
kable(df.tsk %>% select(subID, diagnosis, adhd.meds.bin) %>% distinct() %>% 
        group_by(diagnosis, adhd.meds.bin) %>% count())

kable(df.tsk %>% select(subID, diagnosis, adhd.meds) %>% distinct() %>% 
        filter(!is.na(adhd.meds)) %>%
        group_by(diagnosis, adhd.meds) %>% count())

# drop the neutral condition for the analysis
df.pal = df.tsk %>%
  filter(expected != "neutral" & !is.na(rt.cor)) %>% droplevels() %>%
  mutate_if(is.character, as.factor) %>%
  ungroup()

# calculate variance of reaction times: 
# no difficulty due to suboptimal posterior fit
df.var = df.tsk %>%
  filter(expected != "neutral") %>% droplevels() %>%
  group_by(subID, diagnosis, phase, expected) %>% 
  summarise(
    totaln = n(),
    valuen = sum(!is.na(rt.cor)),
    rt.var = sd(rt.cor, na.rm = T)
  ) %>%
  mutate(
    perc = valuen/totaln
  ) %>% filter(perc >= 2/3) %>%
  mutate_if(is.character, as.factor) %>% ungroup()

# set and print the contrasts
contrasts(df.pal$diagnosis) = contr.sum(3)
contrasts(df.pal$diagnosis)
contrasts(df.pal$expected) = contr.sum(2)
contrasts(df.pal$expected)
contrasts(df.pal$phase) = contr.sum(3)
contrasts(df.pal$phase)
contrasts(df.pal$difficulty) = contr.sum(3)
contrasts(df.pal$difficulty)

contrasts(df.var$diagnosis) = contr.sum(3)
contrasts(df.var$diagnosis)
contrasts(df.var$expected) = contr.sum(2)
contrasts(df.var$expected)
contrasts(df.var$phase) = contr.sum(3)
contrasts(df.var$phase)

# print final group comparisons for the paper into a Word document
read_docx() %>%
    body_add_table(ls.demo$df.demo %>% arrange(measurement) %>% 
                         mutate(bf.log = 
                                  if_else(
                                    bf.log > log(3), 
                                    sprintf("%.3f*", bf.log),
                                    sprintf("%.3f", bf.log)))) %>%
    print(target = "PAL-ADHD_demo.docx")

read_docx() %>%
    body_add_table(ls.demo$df.post %>% arrange(measurement) %>% 
                         mutate(bf.log = 
                                  if_else(
                                    bf.log > log(3), 
                                    sprintf("%.3f*", bf.log),
                                    sprintf("%.3f", bf.log)))) %>%
    print(target = "PAL-ADHD_post.docx")

```

# Reaction time variance

In the preregistration, we noted the following population-level effects for the model of the reaction time variances: group, expectancy, phase and difficulty. However, the posterior fit for this model was suboptimal, therefore, we dropped the predictor difficulty. 

## Model setup

```{r model_var}

# figure out slopes for subjects
kable(head(df.var %>% count(subID, expected)))
kable(head(df.var %>% count(subID, phase)))
kable(head(df.var %>% count(subID, expected, phase)))

# model formula
f.var = brms::bf(rt.var ~ diagnosis * expected * phase +
                   (expected + phase | subID)
                 )

# set weakly informative priors
priors = c(
  prior(normal(4.5,  0.50),  class = Intercept),
  prior(normal(0,    0.50),  class = sigma),
  prior(normal(0.15, 0.15),  class = sd),
  prior(normal(0,    0.25),  class = b),
  prior(lkj(2),              class = cor)
)

```

## Posterior predictive checks

As the next step, we fit the model, check whether there are divergence or rhat issues, and then check whether the chains have converged.

```{r postpc_var1, fig.height=9, message=T}

# fit the final model
m.var = brm(f.var, seed = 4646,
            df.var, prior = priors,
            family = lognormal,
            iter = iter, warmup = warm,
            backend = "cmdstanr", threads = threading(8),
            file = file.path(brms_dir, "m_var"),
            save_pars = save_pars(all = TRUE)
            )
rstan::check_hmc_diagnostics(m.var$fit)

# check that rhats are below 1.01
sum(brms::rhat(m.var) >= 1.01, na.rm = T)

# check the trace plots
post.draws = as_draws_df(m.var)
mcmc_trace(post.draws, regex_pars = "^b_",
           facet_args = list(ncol = 4)) +
  scale_x_continuous(breaks=scales::pretty_breaks(n = 3)) +
  scale_y_continuous(breaks=scales::pretty_breaks(n = 3))

```

This model has no pathological behaviour with E-BFMI, no divergent sample and no rhats that are higher or equal to 1.01. Therefore, we go ahead and perform our posterior predictive checks. 

```{r postpc_var2, fig.height=9}

# get posterior predictions
post.pred = posterior_predict(m.var, ndraws = nsim)

# check the fit of the predicted data compared to the real data
p1 = pp_check(m.var, ndraws = nsim) + 
  theme_bw() + theme(legend.position = "none") + xlim(0, 1000)

# distributions of means compared to the real values per group
p2 = ppc_stat_grouped(df.var$rt.var, post.pred, df.var$diagnosis) + 
  theme_bw() + theme(legend.position = "none")

# distributions of means compared to the real values per phase
p3 = ppc_stat_grouped(df.var$rt.var, post.pred, df.var$phase) + 
  theme_bw() + theme(legend.position = "none")

# distributions of means compared to the real values per expected
p4 = ppc_stat_grouped(df.var$rt.var, post.pred, df.var$expected) + 
  theme_bw() + theme(legend.position = "none")

p = ggarrange(p1, p2, p3, p4,
          nrow = 4, ncol = 1, labels = "AUTO")
annotate_figure(p, top = text_grob("Posterior predictive checks: RT variance", 
                                   face = "bold", size = 14))

```

The simulated data based on the model (light blue) fits well with the real data (dark blue), both for the overall distribution (A) and the aggregated scores for the diagnostic groups (B). 

## Inferences

Now that we are convinced that we can trust our model, we have a look at its estimate and use the hypothesis function to assess our hypotheses and perform explorative tests. 

```{r inf_var, fig.height=9}

# print a summary
summary(m.var)

# get the estimates and compute group comparisons
df.m.var = post.draws %>% 
  select(starts_with("b_")) %>%
  mutate(
    b_COMP         = - b_diagnosis1 - b_diagnosis2,
    b_postvolatile = - b_phase1 - b_phase2,
    BOTH           = b_Intercept + b_diagnosis2,
    ADHD           = b_Intercept + b_diagnosis1,
    COMP           = b_Intercept + b_COMP,
    `ADHD-BOTH`    = ADHD - BOTH,
    `ADHD-COMP`    = ADHD - COMP,
    `BOTH-COMP`    = BOTH - COMP
  )

# plot the posterior distributions
df.m.var %>%
  select(starts_with("b_")) %>%
  pivot_longer(cols = starts_with("b_"), names_to = "coef", values_to = "estimate") %>%
  subset(!startsWith(coef, "b_Int")) %>%
  mutate(
    coef = substr(coef, 3, nchar(coef)),
    coef = str_replace_all(coef, ":", " x "),
    coef = str_replace_all(coef, "diagnosis1", "ADHD"),
    coef = str_replace_all(coef, "diagnosis2", "ADHD+ASD"),
    coef = str_replace_all(coef, "expected1", "expected"),
    coef = str_replace_all(coef, "expected2", "unexpected"),
    coef = str_replace_all(coef, "phase1", "prevolatile"),
    coef = str_replace_all(coef, "phase2", "volatile"),
    coef = fct_reorder(coef, desc(estimate))
  ) %>% 
  group_by(coef) %>%
  mutate(
    cred = case_when(
      (mean(estimate) < 0 & quantile(estimate, probs = 0.975) < 0) |
        (mean(estimate) > 0 & quantile(estimate, probs = 0.025) > 0) ~ "credible",
      T ~ "not credible"
    )
  ) %>% ungroup() %>%
  ggplot(aes(x = estimate, y = coef, fill = cred)) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  ggdist::stat_halfeye(alpha = 0.7) + ylab(NULL) + theme_bw() +
  scale_fill_manual(values = c(c_dark, c_light)) + theme(legend.position = "none")

# H1a: COMP < ADHD
h1a = hypothesis(m.var, "0 < 2*diagnosis1 + diagnosis2")
h1a

## Exploration

# COMP < BOTH
e1 = hypothesis(m.var, "0 < diagnosis1 + 2*diagnosis2", alpha = 0.025)
e1

# ADHD vs. BOTH
e2 = hypothesis(m.var, "diagnosis1 > diagnosis2", alpha = 0.025)
e2

## Exploration of possible transition effects

## create design matrix to figure out how to set comparisons for hypotheses
df.des = cbind(df.var, 
               model.matrix(~ diagnosis * expected * phase, 
                            data = df.var)) %>%
  select(-subID, -rt.var, -perc, -totaln, -valuen) %>% distinct()

# COMP(vol-pre) > ADHD(vol-pre)
as.data.frame(t(df.des %>% 
    ungroup() %>%
    filter((diagnosis == "ADHD" | diagnosis == "COMP") & 
             ((phase == "volatile") | (phase == "prevolatile"))) %>%
    group_by(diagnosis, phase) %>%
    summarise(across(where(is.numeric), ~ mean(.x))) %>%
    group_by(diagnosis) %>%
    summarise(across(where(is.numeric), ~ diff(.x))) %>%
    select(where(is.numeric)) %>%
    map_df(~ diff(.x)))) %>%
  filter(V1 != 0)
e3 = hypothesis(m.var, "0 < 2*diagnosis1:phase1 + diagnosis2:phase1 
                - (2*diagnosis1:phase2 + diagnosis2:phase2)", 
                alpha = 0.025)
e3

# COMP(vol-pre) > BOTH(vol-pre)
e4 = hypothesis(m.var, "0 < diagnosis1:phase1 + 2*diagnosis2:phase1
                - (diagnosis1:phase2 + 2*diagnosis2:phase2)", 
                alpha = 0.025)
e4

# COMP(post-vol) < ADHD(post-vol)
as.data.frame(t(df.des %>% 
    ungroup() %>%
    filter((diagnosis == "ADHD" | diagnosis == "COMP") & 
             ((phase == "volatile") | (phase == "postvolatile"))) %>%
    group_by(diagnosis, phase) %>%
    summarise(across(where(is.numeric), ~ mean(.x))) %>%
    group_by(diagnosis) %>%
    summarise(across(where(is.numeric), ~ diff(.x))) %>% # post - volatile
    select(where(is.numeric)) %>%
    map_df(~ diff(.x)))) %>% # COMP - ADHD
  filter(V1 != 0) 
e5 = hypothesis(m.var, "0 > 2*diagnosis1:phase1 + diagnosis2:phase1 +
                + 2*(2*diagnosis1:phase2 + diagnosis2:phase2)", 
               alpha = 0.025)
e5

# COMP(post-vol) < BOTH(post-vol)
e6 = hypothesis(m.var, "0 > diagnosis1:phase1 + 2*(diagnosis2:phase1 +
                + diagnosis1:phase2 + 2*diagnosis2:phase2)", 
                alpha = 0.025)
e6

## extract predicted differences in ms instead of log data
df.new = df.var %>% 
  select(diagnosis, phase, expected) %>% 
  distinct() %>%
  mutate(
    condition = paste(diagnosis, phase, expected, sep = "_")
  )
df.ms = as.data.frame(
  fitted(m.var, summary = F, 
               newdata = df.new %>% select(diagnosis, phase, expected), 
               re_formula = NA))
colnames(df.ms) = df.new$condition

# calculate our difference columns
df.ms = df.ms %>%
  mutate(
    COMP          = rowMeans(across(matches("COMP_.*"))),
    ADHD          = rowMeans(across(matches("ADHD_.*"))),
    BOTH          = rowMeans(across(matches("BOTH_.*"))),
    h1a_COMPvADHD = ADHD - COMP,
    e_COMPvBOTH   = BOTH - COMP,
    e_BOTHvADHD   = BOTH - ADHD,
    COMPvol       = rowMeans(across(matches("COMP_volatile.*"))),
    COMPpre       = rowMeans(across(matches("COMP_prevolatile.*"))),
    COMPpost      = rowMeans(across(matches("COMP_postvolatile.*"))),
    ADHDvol       = rowMeans(across(matches("ADHD_volatile.*"))),
    ADHDpre       = rowMeans(across(matches("ADHD_prevolatile.*"))),
    ADHDpost      = rowMeans(across(matches("ADHD_postvolatile.*"))),
    BOTHvol       = rowMeans(across(matches("BOTH_volatile.*"))),
    BOTHpre       = rowMeans(across(matches("BOTH_prevolatile.*"))),
    BOTHpost      = rowMeans(across(matches("BOTH_postvolatile.*"))),
    e_volvpreCOMPvADHD   = (COMPvol  - COMPpre) - (ADHDvol  - ADHDpre),
    e_postvvolCOMPvADHD  = (COMPpost - COMPvol) - (ADHDpost - ADHDvol),
    e_volvpreCOMPvBOTH   = (COMPvol  - COMPpre) - (BOTHvol  - BOTHpre),
    e_postvvolCOMPvBOTH  = (COMPpost - COMPvol) - (BOTHpost - BOTHvol),
    e_volvpreBOTHvADHD   = (BOTHvol  - BOTHpre) - (ADHDvol  - ADHDpre),
    e_postvvolBOTHvADHD  = (BOTHpost - BOTHvol) - (ADHDpost - ADHDvol)
    )

kable(df.ms %>% 
  select(starts_with("e") | starts_with("h")) %>% 
  summarise_all(list(lower.ci = lower_ci, mean = mean, upper.ci = upper_ci)) %>%
  t %>% 
  as.data.frame %>% 
  rownames_to_column() %>%
  separate(rowname, into = c("id", "comparison", "stat"), sep = "_") %>%
  pivot_wider(names_from = stat, values_from = V1) %>%
  arrange(comparison), digits = 2)

# calculate effect sizes
df.effect = post.draws %>%
  mutate(across(starts_with("sd")|starts_with("sigma"), ~.^2)) %>%
  mutate(
    sumvar = sqrt(rowSums(select(., starts_with("sd")|starts_with("sigma")))),
    h1a    = (2*b_diagnosis1 + b_diagnosis2) / sumvar,
    e1     = (b_diagnosis1 + 2*b_diagnosis2) / sumvar,
    e2     = (b_diagnosis1 - b_diagnosis2) / sumvar,
    e3     = (`b_diagnosis1:phase1` + 2*`b_diagnosis2:phase1` - 
                `b_diagnosis1:phase2` - 2*`b_diagnosis2:phase2`) / sumvar,
    e4     = (2*`b_diagnosis1:phase1` + `b_diagnosis2:phase1` - 
                2*`b_diagnosis1:phase2` - `b_diagnosis2:phase2`) / sumvar
  )

kable(df.effect %>% select(starts_with("e")|starts_with("h")) %>%
        pivot_longer(cols = everything(), values_to = "estimate") %>%
        group_by(name) %>%
        summarise(
          ci.lo = lower_ci(estimate),
          mean  = mean(estimate),
          ci.hi = upper_ci(estimate),
          interpret = interpret_cohens_d(mean)
        ), digits = 3
)

```

h1a: *estimate* = `r round(h1a$hypothesis$Estimate,2)` [`r round(h1a$hypothesis$CI.Lower,2)`, `r round(h1a$hypothesis$CI.Upper,2)`], *posterior probability* = `r round(h1a$hypothesis$Post.Prob*100,2)`%

e1 COMP < BOTH: *estimate* = `r round(e1$hypothesis$Estimate,2)` [`r round(e1$hypothesis$CI.Lower,2)`, `r round(e1$hypothesis$CI.Upper,2)`], *posterior probability* = `r round(e1$hypothesis$Post.Prob*100,2)`%

e3 COMP(vol-pre) > ADHD(vol-pre): *estimate* = `r round(e3$hypothesis$Estimate,2)` [`r round(e3$hypothesis$CI.Lower,2)`, `r round(e3$hypothesis$CI.Upper,2)`], *posterior probability* = `r round(e3$hypothesis$Post.Prob*100,2)`%

e4 COMP(vol-pre) > BOTH(vol-pre): *estimate* = `r round(e4$hypothesis$Estimate,2)` [`r round(e4$hypothesis$CI.Lower,2)`, `r round(e4$hypothesis$CI.Upper,2)`], *posterior probability* = `r round(e4$hypothesis$Post.Prob*100,2)`%

## Plots

```{r plot_var, fig.height=6}

# rain cloud plot
df.var %>%
  mutate(
    group = case_when(
      diagnosis == "BOTH" ~ "ADHD+ASD",
      T ~ diagnosis
    )
  ) %>%
  ggplot(aes(expected, rt.var, fill = group, colour = group)) + #
  geom_rain(rain.side = 'r',
boxplot.args = list(color = "black", outlier.shape = NA, show.legend = FALSE, alpha = .8),
violin.args = list(color = "black", outlier.shape = NA, alpha = .8),
boxplot.args.pos = list(
  position = ggpp::position_dodgenudge(x = 0, width = 0.3), width = 0.3
),
point.args = list(show_guide = FALSE, alpha = .5),
violin.args.pos = list(
  width = 0.6, position = position_nudge(x = 0.16)),
point.args.pos = list(position = ggpp::position_dodgenudge(x = -0.25, width = 0.1))) +
  scale_fill_manual(values = col.grp) +
  scale_color_manual(values = col.grp) +
  #ylim(0, 1) +
  labs(title = "", x = "", y = "variance (ms)") +
  facet_wrap(. ~ phase) +
  theme_bw() + 
  theme(legend.position = "bottom", plot.title = element_blank(), 
        legend.direction = "horizontal", text = element_text(size = 15),
        legend.title = element_blank())

ggsave("plots/FigRTvar.svg", units = "cm", width = 27, height = 9)

# plot transition effects specifically
df.var %>% 
  group_by(diagnosis, phase) %>%
  summarise(
    var.mn = mean(rt.var),
    var.se = sd(rt.var) / sqrt(n())
  ) %>%
  ggplot(aes(y = var.mn, x = phase, 
             group = diagnosis, colour = diagnosis)) +
  geom_line(position = position_dodge(0.4), linewidth = 1) +
  geom_errorbar(aes(ymin = var.mn - var.se, 
                    ymax = var.mn + var.se), linewidth = 1, width = 0.5, 
                position = position_dodge(0.4)) + 
  geom_point(position = position_dodge(0.4), size = 5) + 
  labs (x = "phase", y = "rt var (ms)") + 
  scale_fill_manual(values = col.grp) +
  scale_color_manual(values = col.grp) +
  theme_bw() + 
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))

```

# Reaction times in correct trials

In the preregistration, we noted the following population-level effects for the model of the reaction time variances: group, expectancy, phase and difficulty; as well as the group-level predictores subject and trials.

## Model setup

```{r model_rt}

# figure out slopes for subjects
kable(head(df.pal %>% count(subID, expected)))
kable(head(df.pal %>% count(subID, phase)))
kable(head(df.pal %>% count(subID, difficulty)))
kable(head(df.pal %>% count(subID, expected, phase)))
kable(head(df.pal %>% count(subID, expected, difficulty)))
kable(head(df.pal %>% count(subID, phase, difficulty)))
kable(head(df.pal %>% count(subID, expected, phase, difficulty)))

# figure out slopes for trls
kable(head(df.pal %>% count(trl, diagnosis)))

# set the formula
f.pal = brms::bf(rt.cor ~ diagnosis * expected * phase * difficulty +
                   (expected * phase * difficulty | subID) + (diagnosis | trl))

# set informed priors based on previous results
priors = c(
  # informative priors based Lawson et al. and Schad, Betancourt & Vasishth (2019)
  prior(normal(6.0,   0.3),   class = Intercept),
  prior(normal(0.0,   0.5),   class = sigma),
  prior(normal(0,     0.1),   class = sd),
  prior(lkj(2),               class = cor),
  prior(normal(100, 100.0),   class = ndt), 
  prior(normal(0.00,  0.04),  class = b)
)

```

## Posterior predictive checks

As the next step, we fit the model, check whether there are divergence or rhat issues, and then check whether the chains have converged. 

```{r postpc_rt, fig.height=9, message=T}

# fit the final model
m.pal = brm(f.pal, seed = 4466,
            df.pal, prior = priors,
            family = shifted_lognormal,
            iter = iter, warmup = warm,
            backend = "cmdstanr", threads = threading(8),
            file = file.path(brms_dir, "m_pal"),
            save_pars = save_pars(all = TRUE)
            )
rstan::check_hmc_diagnostics(m.pal$fit)

# check that rhats are below 1.01
sum(brms::rhat(m.pal) >= 1.01, na.rm = T)

# check the trace plots
post.draws = as_draws_df(m.pal)
mcmc_trace(post.draws, regex_pars = "^b_",
           facet_args = list(ncol = 6)) +
  scale_x_continuous(breaks=scales::pretty_breaks(n = 3)) +
  scale_y_continuous(breaks=scales::pretty_breaks(n = 3))

```

This model has no pathological behaviour with E-BFMI, no divergent sample and no rhats that are higher or equal to 1.01. Therefore, we go ahead and perform our posterior predictive checks. 

```{r postpc_rt2, fig.height=6}

# get posterior predictions
post.pred = posterior_predict(m.pal, ndraws = nsim)

# check the fit of the predicted data compared to the real data
p1 = pp_check(m.pal, ndraws = nsim) + 
  theme_bw() + theme(legend.position = "none") + xlim(0, 1500)

# distributions of means compared to the real values per group or conditions
p2 = ppc_stat_grouped(df.pal$rt.cor, 
                      post.pred, 
                      df.pal$diagnosis) + 
  theme_bw() + theme(legend.position = "none")
p3 = ppc_stat_grouped(df.pal$rt.cor, 
                      post.pred, 
                      df.pal$expected) + 
  theme_bw() + theme(legend.position = "none")
p4 = ppc_stat_grouped(df.pal$rt.cor, 
                      post.pred, 
                      df.pal$phase) + 
  theme_bw() + theme(legend.position = "none")
p5 = ppc_stat_grouped(df.pal$rt.cor, 
                      post.pred, 
                      df.pal$difficulty) + 
  theme_bw() + theme(legend.position = "none")

p = ggarrange(p1, p2, p3, p4, p5, 
          nrow = 5, ncol = 1, labels = "AUTO")
annotate_figure(p, top = text_grob("Posterior predictive checks: reaction times", 
                                   face = "bold", size = 14))

```

The simulated data based on the model fits well with the real data, although there are slight deviations specifically for the predictor expectedness. However, we judge this to be acceptable.  

## Inferences

Now that we are convinced that we can trust our model, we have a look at its estimate and use the hypothesis function to assess our hypotheses and perform explorative tests. 

```{r inf_rt, fig.height=18}

# print a summary
summary(m.pal)

# get the estimates and compute group comparisons
df.m.pal = post.draws %>% 
  select(starts_with("b_")) %>%
  mutate(
    b_postvolatile = - b_phase1 - b_phase2,
    b_difficult    = - b_difficulty1 - b_difficulty2
  )

# plot the posterior distributions
df.m.pal %>%
  pivot_longer(cols = starts_with("b_"), names_to = "coef", values_to = "estimate") %>%
  subset(!startsWith(coef, "b_Int")) %>%
  mutate(
    coef = substr(coef, 3, nchar(coef)),
    coef = str_replace_all(coef, ":", " x "),
    coef = str_replace_all(coef, "diagnosis1", "ADHD"),
    coef = str_replace_all(coef, "diagnosis2", "ADHD+ASD"),
    coef = str_replace_all(coef, "expected1", "expected"),
    coef = str_replace_all(coef, "expected2", "unexpected"),
    coef = str_replace_all(coef, "phase1", "prevolatile"),
    coef = str_replace_all(coef, "phase2", "volatile"),
    coef = str_replace_all(coef, "difficulty1", "easy"),
    coef = str_replace_all(coef, "difficulty2", "medium"),
    coef = fct_reorder(coef, desc(estimate))
  ) %>% 
  group_by(coef) %>%
  mutate(
    cred = case_when(
      (mean(estimate) < 0 & quantile(estimate, probs = 0.975) < 0) |
        (mean(estimate) > 0 & quantile(estimate, probs = 0.025) > 0) ~ "credible",
      T ~ "not credible"
    )
  ) %>% ungroup() %>%
  ggplot(aes(x = estimate, y = coef, fill = cred)) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  ggdist::stat_halfeye(alpha = 0.7) + ylab(NULL) + theme_bw() +
  scale_fill_manual(values = c(c_dark, c_light)) + theme(legend.position = "none")

## create design matrix to figure out how to set comparisons for hypotheses
df.des = cbind(df.pal %>% select(diagnosis, expected, phase, difficulty), 
               model.matrix(~ diagnosis * expected * phase * difficulty, 
                            data = df.pal)) %>% distinct()

# H1c: COMP(unexp-exp) != ADHD(unexp-exp)
as.data.frame(t(df.des %>% 
    filter(diagnosis != "BOTH") %>%
    group_by(diagnosis, expected) %>%
    summarise(across(where(is.numeric), ~ mean(.x))) %>%
    group_by(diagnosis) %>%
    summarise(across(where(is.numeric), ~ diff(.x))) %>% # unexpected - expected
    select(where(is.numeric)) %>%
    map_df(~ diff(.x)))) %>%
  filter(V1 != 0) # COMP - ADHD

h1c = hypothesis(m.pal, "0 > 2*diagnosis1:expected1 + diagnosis2:expected1")
h1c

# COMP(unexp-exp) != BOTH(unexp-exp)
e.both = hypothesis(m.pal, "0 > diagnosis1:expected1 + 2*diagnosis2:expected1", alpha = 0.025)
e.both

# ADHD(unexp-exp) != BOTH(unexp-exp)
e.clin = hypothesis(m.pal, "diagnosis1:expected1 < diagnosis2:expected1", alpha = 0.025)
e.clin

## exploration: task effects

# expected versus unexpected
e1 = hypothesis(m.pal, "0 > 2*expected1", alpha = 0.025)
e1

# volatile versus prevolatile
e2 = hypothesis(m.pal, "0 < phase1 - phase2", alpha = 0.025)
e2

# volatile versus postvolatile
e3 = hypothesis(m.pal, "0 < phase1 + 2*phase2", alpha = 0.025)
e3

# difficult versus medium
e4 = hypothesis(m.pal, "0 < -difficulty1 - 2*difficulty2", alpha = 0.025)
e4

# medium versus easy
e5 = hypothesis(m.pal, "0 < -difficulty1 + difficulty2", alpha = 0.025)
e5

# difficult versus easy
e6 = hypothesis(m.pal, "0 < -2*difficulty1 - difficulty2", alpha = 0.025)
e6

## extract predicted differences in ms instead of log data
df.new = df.pal %>% 
  select(diagnosis, phase, expected, difficulty) %>% 
  distinct() %>%
  mutate(
    condition = paste(diagnosis, phase, expected, difficulty, sep = "_")
  )
df.ms = as.data.frame(
  fitted(m.pal, summary = F, 
               newdata = df.new %>% select(diagnosis, phase, expected, difficulty), 
               re_formula = NA))
colnames(df.ms) = df.new$condition

# calculate our difference columns
df.ms = df.ms %>%
  mutate(
    COMP_expected     = rowMeans(across(matches("COMP_.*_expected_.*"))),
    COMP_unexpected   = rowMeans(across(matches("COMP_.*_unexpected_.*"))),
    ADHD_expected      = rowMeans(across(matches("ADHD_.*_expected_.*"))),
    ADHD_unexpected    = rowMeans(across(matches("ADHD_.*_unexpected_.*"))),
    BOTH_expected      = rowMeans(across(matches("BOTH_.*_expected_.*"))),
    BOTH_unexpected    = rowMeans(across(matches("BOTH_.*_unexpected_.*"))),
    COMP_unexp_exp    = COMP_unexpected - COMP_expected,
    ADHD_unexp_exp     = ADHD_unexpected  - ADHD_expected,
    BOTH_unexp_exp     = BOTH_unexpected  - BOTH_expected,
    `h1c_COMP-ADHD unexpected-expected`  = ADHD_unexp_exp - COMP_unexp_exp,
    `e_COMP-BOTH unexpected-expected`  = BOTH_unexp_exp - COMP_unexp_exp,
    `e_ADHD-BOTH unexpected-expected`  = ADHD_unexp_exp - BOTH_unexp_exp,
    `e1_unexpected-expected`           = rowMeans(across(matches(".*_unexpected_.*"))) - 
      rowMeans(across(matches(".*_expected_.*"))),
    `e2_volatile-prevolatile`       = rowMeans(across(matches(".*_volatile_.*"))) - 
      rowMeans(across(matches(".*_prevolatile_.*"))),
    `e3_volatile-postvolatile`      = rowMeans(across(matches(".*_volatile_.*"))) - 
      rowMeans(across(matches(".*_postvolatile_.*"))),
    `e4_difficult-medium`      = rowMeans(across(matches(".*_difficult"))) - 
      rowMeans(across(matches(".*_medium"))),
    `e5_medium-easy`      = rowMeans(across(matches(".*_medium"))) - 
      rowMeans(across(matches(".*_easy"))),
    `e6_difficult-easy`     = rowMeans(across(matches(".*_difficult"))) - 
      rowMeans(across(matches(".*_easy")))
  )

kable(df.ms %>% 
  select(starts_with("e") | starts_with("h")) %>% 
  summarise_all(list(lower.ci = lower_ci, mean = mean, upper.ci = upper_ci)) %>%
  t %>% 
  as.data.frame %>% 
  rownames_to_column() %>%
  separate(rowname, into = c("id", "comparison", "stat"), sep = "_") %>%
  pivot_wider(names_from = stat, values_from = V1) %>%
  arrange(comparison), digits = 2)

equivalence_test(df.ms %>% select(`h1c_COMP-ADHD unexpected-expected`,
                                  `e_COMP-BOTH unexpected-expected`,
                                  `e_ADHD-BOTH unexpected-expected`),
                 range = rope_range(m.pal))

# calculate effect sizes
df.effect = post.draws %>%
  mutate(across(starts_with("sd")|starts_with("sigma"), ~.^2)) %>%
  mutate(
    sumvar = sqrt(rowSums(select(., starts_with("sd")|starts_with("sigma")))),
    h1c    = -(2*`b_diagnosis1:expected1` + `b_diagnosis2:expected1`) / sumvar,
    e1     = -(2*b_expected1) / sumvar,
    e6     = -(2*`b_difficulty1` + `b_difficulty2`) / sumvar
  )

kable(df.effect %>% select(starts_with("e")|starts_with("h")) %>%
        pivot_longer(cols = everything(), values_to = "estimate") %>%
        group_by(name) %>%
        summarise(
          ci.lo = lower_ci(estimate),
          mean  = mean(estimate),
          ci.hi = upper_ci(estimate),
          interpret = interpret_cohens_d(mean)
        ), digits = 3
)
  

```


h1c: *estimate* = `r round(h1c$hypothesis$Estimate,2)` [`r round(h1c$hypothesis$CI.Lower,2)`, `r round(h1c$hypothesis$CI.Upper,2)`], *posterior probability* = `r round(h1c$hypothesis$Post.Prob*100,2)`%

## Plots

```{r plot_rt, fig.height=10}

# rain cloud plot including all factors

df.pal %>%
  ggplot(aes(diagnosis, rt.cor, fill = expected, colour = expected)) + #
  geom_rain(rain.side = 'r',
boxplot.args = list(color = "black", outlier.shape = NA, show.legend = FALSE, alpha = .8),
violin.args = list(color = "black", outlier.shape = NA, alpha = .8),
boxplot.args.pos = list(
  position = ggpp::position_dodgenudge(x = 0, width = 0.3), width = 0.3
),
point.args = list(show_guide = FALSE, alpha = .5),
violin.args.pos = list(
  width = 0.6, position = position_nudge(x = 0.16)),
point.args.pos = list(position = ggpp::position_dodgenudge(x = -0.25, width = 0.1))) +
  scale_fill_manual(values = col.cnd) +
  scale_color_manual(values = col.cnd) +
  #ylim(0, 1) +
  labs(title = "Reaction times per subject", x = "", y = "rt (ms)") +
  facet_wrap(difficulty ~ phase) +
  theme_bw() + 
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5), 
        legend.direction = "horizontal", text = element_text(size = 15))

```

```{r plot_rt2, fig.height=4}

# rain cloud plot excluding difficulty
df.pal %>%
  group_by(subID, diagnosis, expected, phase) %>%
  summarise(
    rt.cor = median(rt.cor, na.rm = T)
  ) %>%
  group_by(subID, diagnosis, phase) %>%
  summarise(
    rt.exp = diff(rt.cor)
  ) %>%
  ggplot(aes(phase, rt.exp, fill = diagnosis, colour = diagnosis)) + #
  geom_rain(rain.side = 'r',
boxplot.args = list(color = "black", outlier.shape = NA, show.legend = FALSE, alpha = .8),
violin.args = list(color = "black", outlier.shape = NA, alpha = .8),
boxplot.args.pos = list(
  position = ggpp::position_dodgenudge(x = 0, width = 0.3), width = 0.3
),
point.args = list(show_guide = FALSE, alpha = .5),
violin.args.pos = list(
  width = 0.6, position = position_nudge(x = 0.16)),
point.args.pos = list(position = ggpp::position_dodgenudge(x = -0.25, width = 0.1))) +
  scale_fill_manual(values = col.grp) +
  scale_color_manual(values = col.grp) +
  #ylim(0, 1) +
  labs(title = "", x = "", y = "rt difference (ms)") +
  theme_bw() + 
  theme(legend.position = "bottom", plot.title = element_blank(), 
        legend.direction = "horizontal", text = element_text(size = 15),
        legend.title = element_blank())

ggsave("plots/FigRT.svg", units = "cm", width = 27, height = 9)

# plot transition effects specifically
df.pal %>% 
  group_by(subID, diagnosis, phase, expected) %>%
  summarise(rt.cor = median(rt.cor, na.rm = T)) %>%
  group_by(diagnosis, phase, expected) %>%
  summarise(
    rt.mn = mean(rt.cor),
    rt.se = sd(rt.cor)
  ) %>%
  ggplot(aes(y = rt.mn, x = phase, 
             group = diagnosis, colour = diagnosis)) +
  geom_line(position = position_dodge(0.4), linewidth = 1) +
  geom_errorbar(aes(ymin = rt.mn - rt.se, 
                    ymax = rt.mn + rt.se), linewidth = 1, width = 0.5, 
                position = position_dodge(0.4)) + 
  geom_point(position = position_dodge(0.4), size = 5) + 
  labs (x = "phase", y = "rt (ms)") + 
  ggtitle("Diagnosis x phase x expected") + 
  facet_wrap(. ~ expected) +
  scale_fill_manual(values = col.grp) +
  scale_color_manual(values = col.grp) +
  theme_bw() + 
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))

```

# Explorative analysis of errors

Last but not least, we are going to explore possible differences with regards to mean accuracies using a Bayesian ANOVA.

```{r acc}

# create accuracies dataframe
df.acc = df.tsk %>% filter(expected != "neutral") %>% droplevels() %>%
  group_by(subID, diagnosis, phase, expected, difficulty) %>%
  summarise(acc = mean(acc)*100)

# check normal distribution
kable(df.tsk %>% group_by(subID, diagnosis, phase, expected, difficulty) %>%
  summarise(acc = mean(acc)*100) %>% 
  group_by(diagnosis, phase, expected, difficulty) %>%
  shapiro_test(acc) %>%
    mutate(
      sig = if_else(p < 0.05, "*", "")
    ))

# rank transform the data
df.acc = df.acc %>%
  ungroup() %>%
  mutate(
    racc   = rank(acc)
    )

# run the ANOVA
if (!file.exists(file.path(brms_dir, "aov_acc.rds"))){
  aov.acc   = anovaBF(racc   ~ diagnosis * phase * expected * difficulty, data = df.acc)
  saveRDS(aov.acc, file = file.path(brms_dir, "aov_acc.rds"))
} else {
  aov.acc = readRDS(file.path(brms_dir, "aov_acc.rds"))
}

bf.acc = extractBF(aov.acc, logbf = T)
kable(head(bf.acc %>% arrange(desc(bf))))

# print overall accuracy rates for all the effects included in the best model
df.acc %>% 
  group_by(diagnosis, difficulty, expected) %>% 
  summarise(mean_accuracy = mean(acc, na.rm = T), 
            sd_accuracy = sd(acc, na.rm = T))

df.acc %>% 
  group_by(diagnosis) %>% 
  summarise(mean_accuracy = mean(acc, na.rm = T), 
            sd_accuracy = sd(acc, na.rm = T))

df.acc %>% 
  group_by(expected) %>% 
  summarise(mean_accuracy = mean(acc, na.rm = T), 
            sd_accuracy = sd(acc, na.rm = T))

df.acc %>% 
  group_by(difficulty) %>% 
  summarise(mean_accuracy = mean(acc, na.rm = T), 
            sd_accuracy = sd(acc, na.rm = T))

```

Accuracies were generally high, with a grand average of `r round(mean(df.acc$acc),2)`% accurate responses across diagnostic groups. Accuracies seems to have differed between diagnostic groups. Let's do some plotting to figure out where our differences lie. 

## Plots

```{r plot_acc, fig.height=10}

# rain cloud plot including all factors

df.acc %>%
  ggplot(aes(expected, acc, fill = diagnosis, colour = diagnosis)) + #
  geom_rain(rain.side = 'r',
boxplot.args = list(color = "black", outlier.shape = NA, show.legend = FALSE, alpha = .8),
violin.args = list(color = "black", outlier.shape = NA, alpha = .8),
boxplot.args.pos = list(
  position = ggpp::position_dodgenudge(x = 0, width = 0.3), width = 0.3
),
point.args = list(show_guide = FALSE, alpha = .5),
violin.args.pos = list(
  width = 0.6, position = position_nudge(x = 0.16)),
point.args.pos = list(position = ggpp::position_dodgenudge(x = -0.25, width = 0.1))) +
  scale_fill_manual(values = col.grp) +
  scale_color_manual(values = col.grp) +
  labs(title = "Accuracies per subject", x = "", y = "accuracy (%)") +
  facet_wrap(phase ~ difficulty) +
  theme_bw() + 
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5), 
        legend.direction = "horizontal", text = element_text(size = 15))

```

```{r plot_acc2, fig.height=4}

# plot as line plot without the phases
df.acc %>%
  group_by(subID, diagnosis, expected, difficulty) %>%
  summarise(
    acc = mean(acc, na.rm = T)
  ) %>% 
  group_by(diagnosis, expected, difficulty) %>%
  summarise(
    acc.mn = mean(acc),
    acc.se = sd(acc)
  ) %>%
  ggplot(aes(y = acc.mn, x = difficulty, group = expected, colour = expected)) +
  geom_line(position = position_dodge(0.4), linewidth = 1) +
  geom_errorbar(aes(ymin = acc.mn - acc.se, 
                    ymax = acc.mn + acc.se), linewidth = 1, width = 0.5, 
                position = position_dodge(0.4)) + 
  geom_point(position = position_dodge(0.4), size = 5) + 
  facet_wrap(. ~ diagnosis) +
  #ylim(75,100) +
  labs (x = "difficulty", y = "accuracy") + 
  scale_fill_manual(values = col.cnd) +
  scale_color_manual(values = col.cnd) +
  theme_bw() + 
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))

```