---
title: "S2: analysis of pupil sizes with brms"
author: "I. S. Plank"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
fontsize: 10pt
#geometry: margin=0.5in
---
  
```{r settings, include=FALSE}

knitr::opts_chunk$set(echo = T, warning = F, message = F, fig.align = 'center', fig.width = 7.4)
ls.packages = c("knitr",# kable
    "ggplot2",          # plots
    "brms",             # Bayesian lmms
    "designr",          # simLMM
    "bridgesampling",   # bridge_sampler
    "tidyverse",        # tibble stuff
    "ggpubr",           # ggarrange
    "ggrain",           # geom_rain
    "bayesplot",        # plots for posterior predictive checks
    "SBC",              # plots for checking computational faithfulness
    "rstatix",          # anova
    "effectsize",       # interpret_bf
    "BayesFactor", 
    "bayestestR"        # equivalence_test
)

lapply(ls.packages, library, character.only=TRUE)

# set cores
options(mc.cores = parallel::detectCores())

# set options for SBC
use_cmdstanr = getOption("SBC.vignettes_cmdstanr", TRUE) # Set to false to use rstan instead
options(brms.backend = "cmdstanr")

# using parallel processing
library(future)
plan(multisession)

# Setup caching of results
brms_dir = "./_brms_models"
if(!dir.exists(brms_dir)) {
  dir.create(brms_dir)
}

# load helper functions
source('../helper/fun_bf-sens.R')

# graph settings 
c_light = "#a9afb2"; c_light_highlight = "#8ea5b2"; c_mid = "#6b98b2" 
c_mid_highlight = "#3585b2"; c_dark = "#0072b2"; c_dark_highlight = "#0058b2" 
c_green = "#009E73"
sz = 1
a = 0.5

# custom colour palette
col.grp = c("#FFC107", "#004D40", "#1E88E5")
col.cnd = c("#5D3A9B", "#E66100")

lower_ci = function(var) {
  unname(quantile(var, probs = 0.025))
}
upper_ci = function(var) {
  unname(quantile(var, probs = 0.975))
}

```

<style type="text/css">
  .main-container {
    max-width: 1100px;
    margin-left: auto;
    margin-right: auto;
  }
</style>
  
# Introduction
  
This R Markdown script analyses pupil size data from adults with ASD and adults without any psychiatric diagnosis from the PAL (probabilistic associative learning) task of the EMBA project. The data was preprocessed and then selected based on a cross-validation procedure before being read into this script. 

## Some general settings

```{r set}

# number of simulations
nsim = 250

# set number of iterations and warmup for models
iter = 6000
warm = 1500

# set the maximum treedepth
max_depth = 15

# set the seed
set.seed(2468)

```

## Package versions

The following packages are used in this RMarkdown file: 
  
```{r lib_versions, echo=F}

print(R.Version()$version.string)

for (package in ls.packages) {
  print(sprintf("%s version %s", package, packageVersion(package)))
}

```

## Introduction

We planned to determine the group-level effect subjects following Barr (2013). For each model, experiment specific priors were set based on previous literature or the task (see comments in the code).

## Preparation

First, we load the data and print how many participants in each group. 

```{r prep}

# load the behavioural data
load("../data/PAL-ADHD_data.RData")

# load the pupil size data and the diagnosis
df = read_csv(file.path('../data', 'ADHD_CV_pup_sum', "CV_data_C(expected, Sum)[S.expected].csv"),
              show_col_types = F) %>%
  merge(., df.tsk %>% select(subID, diagnosis) %>% distinct(),
        all = T) %>%
  filter(diagnosis != "ASD") %>%
  drop_na() %>% droplevels() %>%
  mutate_if(is.character, as.factor)

# print the number of participants per group
kable(df %>% select(subID, diagnosis) %>% distinct() %>% group_by(diagnosis) %>% count())

# print the contrasts
contrasts(df$diagnosis) = contr.sum(3)
contrasts(df$diagnosis)
contrasts(df$expected)  = contr.sum(2)
contrasts(df$expected)

```

# Pupil size extracted using CV

## Model setup

```{r model_pup}

# model formula
f.pup = brms::bf(rel_pupil ~ diagnosis * expected + rts +
                   (1 | subID)
                 )

# set weakly informative priors
priors = c(
  prior(normal(0,    0.10),  class = Intercept),
  prior(normal(0.15, 0.15),  class = sigma),
  prior(normal(0.15, 0.15),  class = sd),
  prior(normal(0,    0.05),  class = b)
)

kable(priors)

```

## Posterior predictive checks

As the next step, we fit the model, check whether there are divergence or rhat issues, and then check whether the chains have converged.

```{r postpc_pup1, fig.height=3, message=T}

# fit the final model
m.pup = brm(f.pup, seed = 6644,
            df, prior = priors,
            iter = iter, warmup = warm,
            init = 0.1, 
            control = list(max_treedepth = max_depth),
            backend = "cmdstanr", threads = threading(2),
            file = file.path(brms_dir, "m_pup"),
            save_pars = save_pars(all = TRUE)
            )
rstan::check_divergences(m.pup$fit)
rstan::check_energy(m.pup$fit)

# check_treedepth takes treedepth 10 which we already increased, 
# so we check this manually
sparams  = rstan::get_sampler_params(m.pup$fit, inc_warmup = FALSE)
exceedtd = do.call(rbind, sparams)[, "treedepth__"] >= max_depth
message(
      sprintf('%s of %s iterations saturated the maximum tree depth of %s (%.3f%%).',
              sum(exceedtd), 
              length(exceedtd), 
              max_depth, 
              100 * sum(exceedtd) / length(exceedtd))
    )

# check that rhats are below 1.01
sum(brms::rhat(m.pup) >= 1.01, na.rm = T)

# check the trace plots
post.draws = as_draws_df(m.pup)
mcmc_trace(post.draws, regex_pars = "^b_",
           facet_args = list(ncol = 5)) +
  scale_x_continuous(breaks=scales::pretty_breaks(n = 3)) +
  scale_y_continuous(breaks=scales::pretty_breaks(n = 3))

```

This model has no pathological behaviour with E-BFMI, no divergent sample and no rhats that are higher or equal to 1.01. Furthermore, there are no iterations where the maximum treedepth is exceeded. Therefore, we go ahead and perform our posterior predictive checks. 

```{r postpc_pup2, fig.height=6}

# get posterior predictions
post.pred = posterior_predict(m.pup, ndraws = nsim)

# check the fit of the predicted data compared to the real data
p1 = pp_check(m.pup, ndraws = nsim) + 
  theme_bw() + theme(legend.position = "none")

# distributions of means compared to the real values per group
p2 = ppc_stat_grouped(df$rel_pupil, post.pred, df$diagnosis) + 
  theme_bw() + theme(legend.position = "none")

# distributions of means compared to the real values per group
p3 = ppc_stat_grouped(df$rel_pupil, post.pred, df$expected) + 
  theme_bw() + theme(legend.position = "none")

p = ggarrange(p1, p2, p3, heights = c(1,2,1),
          nrow = 3, ncol = 1, labels = "AUTO")
annotate_figure(p, top = text_grob("Posterior predictive checks: pupil sizes",
                                   face = "bold", size = 14))

```

The simulated data based on the model fits reasonably well with the real data. 

## Inferences

Now that we are convinced that we can trust our model, we have a look at its estimate and use the hypothesis function to assess our hypothesis and explore the data. 

```{r inf_pup, fig.height=6}

# print a summary
summary(m.pup)

# plot the posterior distributions
post.draws %>% 
  select(starts_with("b_"), -b_rts) %>%
  pivot_longer(cols = starts_with("b_"), names_to = "coef", values_to = "estimate") %>%
  subset(!startsWith(coef, "b_Int")) %>%
  mutate(
    coef = substr(coef, 3, nchar(coef)),
    coef = str_replace_all(coef, ":", " x "),
    coef = str_replace_all(coef, "diagnosis1", "ADHD"),
    coef = str_replace_all(coef, "diagnosis2", "BOTH"),
    coef = str_replace_all(coef, "expected1", "expected"),
    coef = fct_reorder(coef, desc(estimate))
  ) %>% 
  group_by(coef) %>%
  mutate(
    cred = case_when(
      (mean(estimate) < 0 & quantile(estimate, probs = 0.975) < 0) |
        (mean(estimate) > 0 & quantile(estimate, probs = 0.025) > 0) ~ "credible",
      T ~ "not credible"
    )
  ) %>% ungroup() %>%
  ggplot(aes(x = estimate, y = coef, fill = cred)) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  ggdist::stat_halfeye(alpha = 0.7) + ylab(NULL) + theme_bw() +
  scale_fill_manual(values = c(credible = c_dark, c_light)) + theme(legend.position = "none")

# get the design matrix to figure out how to set the contrasts
df.des = cbind(df, model.matrix(~ diagnosis * expected, data = df)) %>%
  select(starts_with("diagn") | starts_with("expect")) %>% distinct() %>%
  relocate(diagnosis, expected)

# explore expectancy effect
e.exp = hypothesis(m.pup, "0 > 2*expected1", alpha = 0.025)
e.exp

# H2b: COMP(unexp - exp) != ADHD(unexp - exp)
t(df.des %>% filter(diagnosis != "BOTH") %>%
    group_by(diagnosis) %>%
    summarise(across(where(is.numeric), ~ diff(.x))) %>% # unexpected - expected
    arrange(diagnosis) %>%
    select(where(is.numeric)) %>%
    map_df(~ diff(.x))) # COMP(unexpected - expected) - ADHD(unexpected - expected)

h2b = hypothesis(m.pup, "0 > (4*diagnosis1:expected1 + 2*diagnosis2:expected1)")
h2b
  
# calculate effect sizes
df.effect = post.draws %>%
  mutate(across(starts_with("sd")|starts_with("sigma"), ~.^2)) %>%
  mutate(
    sumvar = sqrt(rowSums(select(., starts_with("sd")|starts_with("sigma")))),
    h2b    = -(4*`b_diagnosis1:expected1` + 2*`b_diagnosis2:expected1`) / sumvar
  )

kable(df.effect %>% select(starts_with("e")|starts_with("h")) %>%
        pivot_longer(cols = everything(), values_to = "estimate") %>%
        group_by(name) %>%
        summarise(
          ci.lo = lower_ci(estimate),
          mean  = mean(estimate),
          ci.hi = upper_ci(estimate),
          interpret = interpret_cohens_d(mean)
        ), digits = 3
)


```

h2b: COMP(unexpected-expected) vs. ADHD(unexpected-expected): *estimate* = `r round(h2b$hypothesis$Estimate,2)` [`r round(h2b$hypothesis$CI.Lower,2)`, `r round(h2b$hypothesis$CI.Upper,2)`], *posterior probability* = `r round(h2b$hypothesis$Post.Prob*100,2)`%

## Plots

```{r plot_pup, fig.height=4}

# rain cloud plot
df %>%
  group_by(subID, diagnosis, expected) %>%
  summarise(
    rel_pupil = mean(rel_pupil)
  ) %>%
  ggplot(aes(diagnosis, rel_pupil, fill = expected, colour = expected)) + #
  geom_rain(rain.side = 'r',
boxplot.args = list(color = "black", outlier.shape = NA, show.legend = FALSE, alpha = .8),
violin.args = list(color = "black", outlier.shape = NA, alpha = .8),
boxplot.args.pos = list(
  position = ggpp::position_dodgenudge(x = 0, width = 0.3), width = 0.3
),
point.args = list(show_guide = FALSE, alpha = .5),
violin.args.pos = list(
  width = 0.6, position = position_nudge(x = 0.16)),
point.args.pos = list(position = ggpp::position_dodgenudge(x = -0.25, width = 0.1))) +
  scale_fill_manual(values = col.cnd) +
  scale_color_manual(values = col.cnd) +
  labs(title = "", x = "", y = "pupil size change (mm)") +
  theme_bw() + 
  theme(legend.position = "bottom", plot.title = element_blank(),
        legend.direction = "horizontal", text = element_text(size = 15), 
        legend.title = element_blank())

ggsave("plots/FigPUP.svg", units = "cm", width = 27, height = 9)


```

